Lesson 5
========================================================

```{r}
library(ggplot2)
library(dplyr)
pf <- read.csv('../lesson4/pseudo_facebook.tsv', sep = '\t')
```


### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) + 
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```

```{r}
pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>%
  ungroup() %>%
  arrange(age)

head(pf.fc_by_age_gender)
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y = mean_friend_count, color = gender), 
              data = subset(pf.fc_by_age_gender, !is.na(gender))) +
  geom_line()
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
#install.packages('reshape2')
library(reshape2)
```

```{r}
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender, 
                                  value.var = 'median_friend_count')
head(pf.fc_by_age_gender.wide)
```

```{r}
pf.fc_by_age_gender.tmp <- subset(pf.fc_by_age_gender[c('age', 'gender', 'median_friend_count')], !is.na(gender))

pf.fc_by_age_gender.wide <- pf.fc_by_age_gender.tmp %>%
  spread(gender, median_friend_count) %>%
  mutate(ratio = male / female)

head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(aes(x = age, y = female/male),
       data=pf.fc_by_age_gender.wide) +
  geom_line() +
  geom_hline(yintercept = 1, alpha = 0.3, linetype = 2)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- floor( 2014 - pf$tenure/365.)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
summary(pf$year_joined)
table(pf$year_joined)
#        (2004, 2009]
#        (2009, 2011]
#        (2011, 2012]
#        (2012, 2014]

pf$year_joined.bucket <- cut(pf$year_joined, c(2004,2009,2011,2012,2014))

```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
table(pf$year_joined.bucket, useNA = 'ifany')

ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), 
            stat = 'summary', 
            fun.y = median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), 
            stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = mean, linetype = 2)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
with(subset(pf, pf$tenure>=1), summary(friend_count/tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate? **0.2205**

What is the maximum friend rate? **417¡£00**

```{r Friendships Initiated}
ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure>=1)) + 
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary', fun.y=mean)

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

p0 <- ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

p1 <- ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

p2 <- ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

p3 <- ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)
library(gridExtra)
grid.arrange(p0, p1, p2, p3, ncol=1)
```

```{r}
ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure>=1)) + 
  geom_smooth(aes(color = year_joined.bucket))
```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv('yogurt.csv')
str(yo)

yo$id <- factor(yo$id)
str(yo)

ggplot(aes(x = price), data=yo) +
  geom_histogram()
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
summary(yo)
length(unique(yo$price))
table(yo$price)

yo <- transform(yo, all.purchass = yo$strawberry +
                  yo$blueberry + yo$pina.colada +
                  yo$plain)
summary(yo$all.purchass)
```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(aes(x = time, y = price), data = yo) +
  geom_jitter(alpha = 1/4, shape = 21,
              fill=I('#F79420'))
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(57680)
sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(x = time, y = price),
       data = subset(yo, id %in% sample.ids)) +
  facet_wrap(~id) +
  geom_line() +
  geom_point(aes(size = all.purchass), pch=1)

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:
```{r}
library(GGally)
theme_set(theme_minimal(20))

set.seed(1836)
pf_subset <- pf[, c(2:15)]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset),1000),])
```


***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

